Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-13 eswc-2012/phdsymp.tex
snip
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@83 43ef1190-90fd-4dd7-a769-d0121f02b80e

Query federation has been an active field for some time, but has until
the advent of the Semantic Web not been used in a highly diverse set
of endpoints, commonly they have been under the control of a single
institution. RDF has a triple-based data model for the Semantic Web,
and SPARQL is a standardized query language to query such data.

@@

In a federated regime, any query is likely to execute queries on
several different SPARQL implementations. Many of those are free
software so the algorithms they use could in principle be analyzed,
but in practice, this would be very time-consuming. Some
implementations may be undocumented or even trade secrets, thus
precluding any such analysis. It is therefore my intention to treat
the individual SPARQL implementations as ``black boxes'' and
exclusively evaluate the performance characteristics by empirical
means, aka ``benchmarking''.

@@

When the difference is small, we must consider the possibility that
the measurements are impacted by random noise. Computer systems are
generally very deterministic systems, but there are nevertheless many
such scenarios: The underlying system may be doing other tasks. The
garbage collection of some languages may trigger at not easily
foreseeable points. The engine may use randomized algorithms, or the
data may be of such a nature that it triggers worst-case behaviour in
situations the query writer cannot foresee. All this contributes to
that we should consider the performance metric, whatever metric is
used, as a \emph{stochastic variable}.

>>>>

@@

@@

In this case, we must consider the possibility that the measurements
are impacted by random noise. Computer systems are generally very
deterministic systems, but there are nevertheless many such scenarios:
The underlying system may be doing other tasks. The garbage collection
of some languages may trigger at not easily foreseeable points. The
engine may use randomized algorithms, or the data may be of such a
nature that it triggers worst-case behaviour in situations the query
writer cannot foresee. All this contributes to that we should consider
the performance metric, whatever metric is used, as a \emph{stochastic
  variable}.
